---
title: "Urn models"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(ggplot2)
library(tidyr)
source("../R/urn_models.R") # the functions
# unfortunately the standard colorset sets red to blue and reverse, so this fixes it.
set_colors <- scale_color_manual(values=c(red="#c02519",blue="#7ca0c0"))
```

These are some models based on the coursera course:
Model Thinking and specificly on [this video](https://www.coursera.org/learn/model-thinking/lecture/ym0VV/urn-models)

Urn models are simple models which can have some complex behavior, the models are like this

* you start with an urn (or bag, or whatever a container)
* there are 2 marbles in the urn (one red, one blue f.i.)
* you draw a marble, look at it and return it

## Simple model, Bernoulli
Nothing more than description above and basic idea of probability classes.
Every draw is independent.

```{r}
urn_bernoulli(iterations = 10)
```
How many of every color would we have if we repeat this proces 100 times?
```{r}
urn_plot(cumulative_sum(urn_bernoulli(iterations = 100)))+
    set_colors+
    ggtitle("Bernoulli")
```

```{r}
urn_plot(cumulative_sum(urn_bernoulli(iterations = 100)))+
    set_colors+
    ggtitle("Bernoulli", subtitle = "another try")
```

Because there will always be 1 red and 1 blue ball, the long run probability
of drawing 'red' will be 50%.

A look over the prob distribution over many repeats:

```{r}
vis_prop_distr(prob_distr_urn(urn_bernoulli,iterations_per_model = 100,repeats = 200))+
    ggtitle("Bernoulli")
```

## Extending the model, Polya model

* you start with an urn (or bag, or whatever a container)
* there are 2 marbles in the urn (one red, one blue f.i.)
* you draw a marble, look at it and return it
* **Add another marble of that color to the bag**


```{r}
urn_polya(iterations = 10)
```
How many of every color would we have if we repeat this proces 100 times?
```{r}
urn_plot(cumulative_sum(urn_polya(iterations = 100)))+
    set_colors+
    ggtitle("Polya model")
```

Is it different with a different start?
```{r}
urn_plot(cumulative_sum(urn_polya(iterations = 100)))+
    set_colors+
    ggtitle("Polya model", subtitle = "another draw")
```

A look over the prob distribution over many repeats:

```{r}
vis_prop_distr(prob_distr_urn(urn_polya,iterations_per_model = 100,repeats = 200))+
    ggtitle("Polya",subtitle = "flat distribution, anything is possible")
```


## Extending the model, Balancing model

* you start with an urn (or bag, or whatever a container)
* there are 2 marbles in the urn (one red, one blue f.i.)
* you draw a marble, look at it and return it
* **Add another marble of a different color to the bag**


```{r}
urn_balancing(iterations = 10)
```
How many of every color would we have if we repeat this proces 100 times?
```{r}
urn_plot(cumulative_sum(urn_balancing(iterations = 100)))+
    set_colors+
    ggtitle("Balancing model")
```

Is it different with a different start?
```{r}
urn_plot(cumulative_sum(urn_balancing(iterations = 100)))+
    set_colors+
    ggtitle("Balancing model", subtitle = "another draw")
```

A look over the prob distribution over many repeats:

```{r}
vis_prop_distr(prob_distr_urn(urn_balancing,iterations_per_model = 100,repeats = 200))+
    ggtitle("Balancing",subtitle = "Prob. moves to equilibrium")
```

## Extending the model, Sway model

* you start with an urn (or bag, or whatever a container)
* there are 2 marbles in the urn (one red, one blue f.i.)
* you draw a marble, look at it and return it
* **Add another marble of that color to the bag**
* **We add 1 balls of color step back (2^{0})**
*  **We add 2 balls of color 2 steps back (2^{1})**
*  **We add 4 balls of color 3 steps back (2^{2})**
*  **etc. **

In this model the past weights more heavily over time.

```{r}
urn_sway(iterations = 10)
```
How many of every color would we have if we repeat this proces 100 times?
```{r}
urn_plot(cumulative_sum(urn_sway(iterations = 100)))+
    set_colors+
    ggtitle("Sway model")
```

Is it different with a different start?
```{r}
urn_plot(cumulative_sum(urn_sway(iterations = 100)))+
    set_colors+
    ggtitle("Sway model", subtitle = "another draw")
```

A look over the prob distribution over many repeats:

```{r}
vis_prop_distr(prob_distr_urn(urn_sway,iterations_per_model = 100,repeats = 200))+
    ggtitle("Sway",subtitle = "flat distribution, anything is possible")
```


## Path dependence and increasing returns, Symbiots urn

```{r}
urn_symbiots(iterations = 10)
```


```{r}
urn_plot(cumulative_sum(urn_symbiots(iterations = 100)))+
    scale_color_manual(values=c(red="#c02519",blue="#7ca0c0", yellow="#fff855",green="#6a7f50"))+
    ggtitle("Symbiot urn model")
```

```{r}
urn_plot(cumulative_sum(urn_symbiots(iterations = 100)))+
    scale_color_manual(values=c(red="#c02519",blue="#7ca0c0", yellow="#fff855",green="#6a7f50"))+
    ggtitle("Symbiot urn model", subtitle = "different")
```

